Performance for the masses: Experiments with A Web Based Architecture to Harness Volunteer Resources for Low Cost Distributed Evolutionary Computation

### 2016-02-03 GECCO/volunteer-clouds-gecco16.tex
Little changes and comments

The fact that the solution of this particular problem was found in a
few seconds and restarts were necessary yielded interesting results,
but we needed to work on a harder problem, that is why created a new
client version that optimized a harder variation of the Rastrigin's function, 
used as a benchmark in evolutionary
algorithms. The set up for this function, and a comparison of the
baseline speed for JavaScript in an standalone experiment and other
languages is shown in \cite{2016arXiv160101607Manom}, where the details of
the implementation have also been extensively described.

Specifically, this experiment was deployed in a virtual private server with 512MB of memory
running Ubuntu OS version 14.04 and hosted in DigitalOcean.com. Unlike previous
experiments where the optimum was expected to be found many times, 
this problem was expected to run constantly for many days finding only
sub-optimal approximations. In order to achieve this, the only change needed 
in the framework was to limit the number of chromosomes kept in the pool. 
Previously, every PUT request which included a new chromosome increased 
the size of the pool. But in a long-running experiment this has to be avoided
in order to not exhaust the available memory. In this experiment the size
of a chromosome was 1000 double precision numbers, so that considering this
and the memory available in the server, the pool was limited 
to 10,000 chromosomes.  In order to keep this size
constant, a random chromosome was pulled out of the pool before inserting a
new one, when the pool was full. For this problem {\sf NodEO}'s {\tt chromosome-float} 
and {\tt classic-ea-float} modules were used. The parameters for the EA 
were tournament size = 3 and  population size = 500, returning exchanging a
chromosome with the server every 100 generations. Again, there were two workers in each window.
As in previous experiments, a few calls to participation were published in social networks.

>>>>

% (Paloma) I would propose as section title "A hard optimization problem: The Shifted Rotated Rastrigin's function"

The fact that the solution of this particular problem was found in a
few seconds and restarts were necessary yielded interesting results, % (Paloma) Where?
but we needed to work on a harder problem, that is why created a new
client version that optimized a harder variation of the Rastrigin's function, 
used as a benchmark in evolutionary
algorithms. The set up for this function, and a comparison of the
baseline speed for JavaScript in an standalone experiment and other
languages is shown in \cite{2016arXiv160101607Manom}, where the details of
the implementation have also been extensively described.

Specifically, this experiment was deployed in a virtual private server with 512MB of memory
running Ubuntu OS version 14.04 and hosted in DigitalOcean.com. Unlike previous
experiments where the optimum was expected to be found many times, 
this problem was expected to constantly run for many days finding only
sub-optimal approximations. In order to achieve this, the only change needed 
in the framework was to limit the number of chromosomes kept in the pool. 
Previously, every PUT request which included a new chromosome increased 
the size of the pool. But in a long-running experiment this has to be avoided
in order to not exhausting the available memory. In this experiment the size
of a chromosome was 1000 double precision numbers, so that considering this
and the memory available in the server, the pool was limited 
to 10,000 chromosomes.  In order to keep this size
constant, a random chromosome was pulled out of the pool before inserting a
new one, when the pool was full. For this problem {\sf NodEO}'s {\tt chromosome-float} 
and {\tt classic-ea-float} modules were used. The parameters for the EA 
were tournament size = 3 and  population size = 500, returning exchanging a
chromosome with the server every 100 generations. Again, there were two workers in each window.
As in previous experiments, a few calls to participation were published in social networks.
